Robot Team III: Scotty Strikes Back

### 2024-04-16 paper/paper.tex
Updated intro text

1. Classification is a common remote sensing and hyperspectral imaging problem
    - Endmember Identification and Spectral Unmixing is sub-problem specific to hyperspectral data

2. HSI Size allows their incorporation onto drones
    - \cite{robot-team-1}
    - \cite{robot-team-2}

3. Provided with in situ measurements, HSI can be used to predict a variety of relevant water quality and water composition parameters. Here we can describe our previous work

4. However, we would like to be able to identify the presence of spectral signatures for which we do not already have an in situ sensor.

5. Unsupervised classification techniques are nice however standard algorithms like k-means clustering require prior knowledge of the number of constituent endmembers and do not, in general, provide any interpretable relationship between classes e.g. how is class 1 more similar to class 2 than to class 10. The Self Organizing map introduced by Kohonen is a popular technique which enables an unsupervised classification of data where classes

- We should make a note about Autoencoders and VAEs here. While they work well in many contexts, they do not provide a relationship between classes and there is no obvious way to choose the size of the encoding

6. "In this paper, we extend our previous work into the unsupervised classification domain..." Here we demonstrate the application of Generative Topographic mapping which addresses these issues by providing a probability distribution of resulting classes.

7. We apply the GTM to datasets collected by our Robot Team specifically showcasing the identification of spectral signatures for Rhodamine Dye, Algae, etc...

8. With the signatures identified, class abundance, and therefore area estimation, can be performed rapidly, in real time, by utilizing the learned signatures with the nromalized spectral similarity score (or other similar metrics).

IDEA: spectral unmixing of water is challenging due to nonlinear effects of scattering by solids and absorption by dissolved chemicals (dyes etc...) this motivates a nonlinear approach to endmember identification, etc...

>>>>

Inland water bodies pose a unique challenge for analysis by remote sensing imagery due to their complex spectral characteristics and small-scale spatial variability; nevertheless, significant effort has gone into the development of techniques for estimating optically-active water quality parameters such as chlorophyll-a, blue-green algae, turbidity, and temperature (citation). Machine learning techniques have proven to be a powerful tool in this domain, facilitating the development of data-driven models which map observed reflectance to parameters of interest. However, these methods rely on the availability of significant volumes of in-situ reference data needed to calibrate models and asses their uncertainty. For example Aurin et al... Similarly, Ross et al developed... The challenge posed by in-situ data collection together with long satellite revisit times limits the utility of remote sensing imagery for real-time applications.

- supervised classification of water quality with XGBoost \cite{ghatkar2019classification}

Unsupervised classification

endmember extraction, spectral unmixing

Hyperspectral imaging has emerged as a pivotal technology in this space. With their increased spectral resolution, hyperspectral imagers (HSIs) capture the complex chemical, biological, environmental characteristics of reflectance spectra. (add references for specific applications of HSIs). Continued development of HSI technology has led to a dramatic reduction in their size enabling their utilize as the payload of small autonomous aerial vehicles (UAVs) (citation). By flying close to the ground UAV-born hyperspectral imagers can produce centimeter-scale imagery while simulatenously limiting the need for complicated atmospheric corrections required for satellite platforms (citation).

In our previous work, we introduced a robot team designed to rapidly characterize water composition in novel environments \cite{robot-team-1}. 
Joining UAV-borne HSI together with collection of in-situ reference data by an autonomous boat dramatically reduces the time required to generate .

By carefully coordinating the simultanous collection of ground-truth water quality data and hyperspectral imagery, this team can rapidly estimate concentrations of optically-active components including colored dissolved organic matter (CDOM), blue-green algae, crude oil, temperature as well

In this paper, we demonstrate how Generative Topographic Mapping (GTM), a Bayesian alternative to the Self Organizing Map, can be utilized to rapidly identify relevant spectral signatures within a scene. Using data collected at site in Montague, North Texas, we demonstrate the ability of this paradigm with two case studies. First, we demonstrate how the GTM can be used to provide a high-resolution unsupervised classification of a water body enabling the rapid identification of relevant regions which can then be used to direct further investigation by reference-sensor equipped boats. Second, we demonstrate the ability to perform endmember extraction and abundance mapping using imagery from a test-release of rhodamine dye. Together with the normalized spectral similarity score, these spectral signatures are identified and their associated area in the scene is determined.
